Ornaments as indicators of social changes and cultural practices in northeast Taiwan before and after European colonial period

### 2019-12-06 analysis/paper/paper.Rmd
added code chunk for c14 dates

Currently the ornaments found in northeast Taiwan in early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metal working, or accessible local raw materials. The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018]. Although there is wide varity of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011]. There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007]. Northeast Taiwan from the late Iron age to early historical period was involved in complex exchange network both in a regional and global scale.

@@

```{r C14-table-data}
C14 <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
knitr::kable(C14,
             caption = "Radiocarbon dates on charcoals from Kiwulan")

```

@@

Another scenario could be that the increasing number and concentrated spatial pattern of ornaments indicate cultural resistance against the European intrusion by using them to show social identity and emphasize the local custom that had existed before European contact. Resistance could be presented in many forms. For example, @Rubertone discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands. Similarly, the frequent presence of ornaments across the dwelling area during the European suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition. During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan. This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or population decrease. It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next. This might suggest that the ornaments found in Chinese period were old item continuously used till then. Further analysis about the source for ornaments from different periods will be helpful to understand their origin and reginal circulation.

@@

# Acknowledgments

>>>>

Currently the ornaments found in northeast Taiwan in early historical period, including glass beads, stone beads, and metal ornaments, were considered as imported materials from other regions due to a lack of evidence of beadmaking waste, metal working, or accessible local raw materials. The chemical composition analysis for glass beads shows high content of lead, coupled with the winding/folding technique, suggest a Chinese beadmaking tradition [@Cheng2008; @Gan2006; @Wang2018]. Although there is wide variety of metal ornaments such as bells, bracelets, rings, and pendants, the common compositions of metal ornaments are brass and copper with a small number made from lead and tin that indicate multiple origins mostly likely from Southeast Asia [@Chen2011]. There is no clear evidence shows the European origin of beads, however, the large amount of golden beads at Kiwulan might be related to the colonial presence of the Europeans through economic activities because similar beads were found at Luzon, northern Philippines, a part of the trading route of the Spanish between 16-19th century [@Wang2007]. Northeast Taiwan from the late Iron age to early historical period was involved in complex exchange network both in a regional and global scale.

@@

```{r C14-table-data}
kw_dates <- readxl::read_excel(here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
table_kw_dates <-
  kw_dates %>% 
  select(-se, -rce)

knitr::kable(table_kw_dates,
             caption = "Radiocarbon dates on charcoals from Kiwulan")
```

```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan. The y-axis represents summed probability and the x-axis is the date in calendar years. Produced with the BChronDensity() function in program R's Bchron package (Parnell et al. 2008). IntCal13 Atmospheric curve."}

# subset dates with error value
kw_dates_with_error <- 
  kw_dates %>% 
  filter(!is.na(se))

# convert rce to numeric
kw_dates_with_error$rce <- as.numeric(kw_dates_with_error$rce)

# make some kind of ID
kw_dates_with_error$id <- with(kw_dates_with_error,
                               paste0(Square,"-",Layer))

# Use Bchron to clibrate
library(Bchron)

kw_dates_with_error_calibrated <- 
  with(kw_dates_with_error,
       BchronCalibrate(
         ages = rce,
         ageSds = se,
         ids = id,
         calCurves = rep('intcal13',
                         length(kw_dates_with_error$rce))
       ))

# have a look at the output
summary(kw_dates_with_error_calibrated)
par(mfrow = c(2, 3))
plot(kw_dates_with_error_calibrated)

# we can combine all the calibrated dates into a single plot, like this
kw_ages_densities <- 
  with(kw_dates_with_error,
       BchronDensity(
         ages = rce,
         ageSds = se,
         calCurves = rep('intcal13',
                         length(kw_dates_with_error$rce))
       ))

#plot it
plot(kw_ages_densities,
     xlab= "Years BC")

# to describe in your text, ?BchronDensity
# "We used the Bchron R package to compute a non-parametric phase model 
# on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
# FIgure X shows the results of this analysis... indicating that the peak period of occupation is around 600 BC"
```

@@

Another scenario could be that the increasing number and concentrated spatial pattern of ornaments indicate cultural resistance against the European intrusion by using them to show social identity and emphasize the local custom that had existed before European contact. Resistance could be presented in many forms. For example, @Rubertone discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands. Similarly, the frequent presence of ornaments across the dwelling area during the European suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition. During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan. This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or population decrease. It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next. This might suggest that the ornaments found in Chinese period were old item continuously used until then. Further analysis about the source for ornaments from different periods will be helpful to understand their origin and reginal circulation.

@@

# Acknowledgements
